Exploring the Demographic Aspects of COVID-19 Cases in Toronto: Are the elderly males more susceptible to COVID-19?

### 2024-01-25 outputs/paper/paper.qmd
Update Paper
Add appendix
Update paper and llm
Add table of contents and discussion section in paper

Add ChatGPT usage in llm file

---
title: "Exploring the Demographic Aspects of COVID-19 Cases in Toronto: Are the elderly males more susceptible to COVID-19?"
author: 
  - Amie Liu
thanks: "Code and data are available at:https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
date: today
date-format: long
abstract: "Examining the age and gender patterns of COVID-19 cases in Toronto provides insights into the susceptibility of different demographic groups to the virus. This paper utilizes the dataset from Open Data Toronto, encompassing the period from January 22, 2020, to January 4, 2024. Contrary to the hypothesis that elderly males are more susceptible, the results reveal a pattern, with females aged 20-29 showing the highest susceptibility, exceeding 40,000 cases. This highlights the importance of understanding gender and age-specific patterns for effective public health interventions."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper is divided into the Data and Limitations sections. @sec-data Data delves into the various sources used for data collection, cleaning, and analyzing processes, leveraging a comprehensive dataset provided by Toronto Public Health on the City of Toronto's Open Data Portal [@covidCases].  In addition, it provides a detailed analysis of the findings from the data. Moving forward, @sec-limitations Limitations reveal potential constraints, including considerations related to occupational exposure, geographical focus, variable measurement, and data accuracy, that might influence the interpretation of the main findings.

@@

The data for this paper was obtained from the City of Toronto's Open Data Portal through the 'opendatatoronto' library [@openDataToronto]. The details of the data cleaning process for the dataset 'COVID-19 Cases in Toronto' [@covidCases] are provided in @sec-appendix Appendix. The data analysis involved using the R programming language [@citeR] along with the 'tidyverse' [@citeTidyverse] 'here' [@citeHere], 'janitor' [@citeJanitor], 'knitr' [@citeKnitr] packages for data collection, cleaning, and analysis. Within the 'tidyverse' package [@citeTidyverse], I used 'readr' [@citeReadr], 'dplyr' [@citeDplyr], 'tidyr' [@citeTidyr], and 'ggplot2' [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating bar charts.

## Covid-19 Cases by Gender

@@

## Covid-19 Cases by Age Groups

@@

## Distribution of Covid-19 Cases Across Genders and Age Groups

@@

# Limitations {#sec-limitations}

@@

## Occupational Exposure

@@

## Geographical focus

@@

## Variable Measurement

@@

## Data Accuracy

@@

Future research could address these limitations by exploring other demographic variables and considering a wider geographical scope. This would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases in diverse populations.

# Appendix {#sec-appendix}

## Data Cleaning

I conducted a data-cleaning procedure, focusing on variable selection, simplifying variable names, and handling missing values. The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender. If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable). The total number of rows has been reduced from 412,122 to 411,700.

```{r}
#| label: tbl-idcasetable
#| tbl-cap: First six rows of COVID-19 cases by age groups and gender
#| echo: false
#| message: false
#| warning: false

>>>>

---
title: "Exploring the Demographic Aspects of COVID-19 Cases in Toronto: Are the elderly males more susceptible to COVID-19?"
author: 
  - Amie Liu
thanks: "Code and data are available at:https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
date: today
date-format: long
abstract: "Examining the age and gender patterns of COVID-19 cases in Toronto provides insights into the susceptibility of different demographic groups to the virus. This paper utilizes the dataset from Open Data Toronto, encompassing the period from January 22, 2020, to January 4, 2024. Contrary to the hypothesis that elderly males are more susceptible, the results reveal a pattern, with females aged 20-29 showing the highest susceptibility, exceeding 40,000 cases. This highlights the importance of understanding gender and age-specific patterns for effective public health interventions."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

This paper is divided into the Data and Limitations sections. @sec-data Data delves into the various sources used for data collection, cleaning, and analyzing processes, leveraging a comprehensive dataset provided by Toronto Public Health on the City of Toronto's Open Data Portal [@covidCases].  In addition, @sec-results Results provides a detailed analysis of the findings from the data. Moving forward, @sec-discussion Discussion provides an understanding of the demographic patterns of COVID-19 cases, highlights key findings, and reveals potential limitations, including considerations related to occupational exposure, geographical focus, variable measurement, and data accuracy, that might influence the interpretation of the main findings.

@@

The data for this paper was obtained from the City of Toronto's Open Data Portal through the 'opendatatoronto' library [@openDataToronto]. The data analysis involved using the R programming language [@citeR] along with the 'tidyverse' [@citeTidyverse] 'here' [@citeHere], 'janitor' [@citeJanitor], 'knitr' [@citeKnitr] packages for data collection, cleaning, and analysis. Within the 'tidyverse' package [@citeTidyverse], I used 'readr' [@citeReadr], 'dplyr' [@citeDplyr], 'tidyr' [@citeTidyr], and 'ggplot2' [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating bar charts.

## COVID-19 Cases Data

The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender. If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable). The total number of rows has been reduced from 412,122 to 411,700.

```{r}
#| label: tbl-idcasetable
#| tbl-cap: Sample of cleaned COVID-19 cases data
#| echo: false
#| message: false
#| warning: false

# Results {#sec-results}

## COVID-19 Cases by Gender

@@

## COVID-19 Cases by Age Groups

@@

## Distribution of COVID-19 Cases Across Genders and Age Groups

@@

# Discussion {#sec-discussion}

## Interpretation of Findings
The examination of COVID-19 cases by gender reveals that females represent 53.4% of the cases in Toronto and males represent 45.7% (see @fig-casebygender), which does not show significant differences between female and male cases. The age-stratified analysis of COVID-19 cases shows that the 20-29 age group is most affected by COVID-19, accounting for 18.83% of the overall cases (see @fig-casebyagegroup). Conversely, older age groups (40+) have lower case numbers. Moreover, the intersectional analysis of gender and age groups shows that most age groups have a higher number of female cases, with exceptions in the 0-19 and 70-79 age groups where male cases are higher than female cases (see @fig-distribution). These findings do not support the hypothetical question: "Are elderly males more susceptible to COVID-19?" Additional demographic variables, such as socioeconomic status and cultural factors, may influence COVID-19 patterns in diverse populations.

## Limitations

@@

### Occupational Exposure

@@

### Geographical focus

@@

### Variable Measurement

@@

### Data Accuracy

@@

## Next Steps

Future research could address these limitations by exploring other demographic variables and considering a wider geographical scope. This exploration would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases across diverse populations. Future research endeavors have the potential to refine public health strategies, tailor interventions to specific population needs, and contribute substantially to the global understanding of the multifaceted nature of the COVID-19 pandemic.
